### 2021-03-12 paper/Paper/Section2-Kits.lagda.tex
Customized agda.sty and started section 2.

\begin{code}[hide]
module Paper.Section2-Kits where

open import Data.List using (List; []; _∷_)

infix 4 _∋_ _⊢_
infixr 5 _⇒_ λx_
infixl 6 _·_
\end{code}

\newpage
\section{Kits}

McBride proposed the kit technique for intrinsically typed formalizations.

\begin{code}
data Type : Set where
  ∅    : Type
  _⇒_  : Type → Type → Type

variable
  t t₁ t₂  : Type
  Γ        : List Type

-- Well-typed DeBruijn variables
data _∋_ : List Type → Type → Set where
  zero  : (t ∷ Γ ∋ t)
  suc   : (Γ ∋ t₁) → (t₂ ∷ Γ ∋ t₁)

-- Well-typed terms
data _⊢_ : List Type → Type → Set where
  `_   : (Γ ∋ t) → (Γ ⊢ t)                    -- Variables
  λx_  : (t₁ ∷ Γ ⊢ t₂) → (Γ ⊢ t₁ ⇒ t₂)        -- Functions
  _·_  : (Γ ⊢ t₁ ⇒ t₂) → (Γ ⊢ t₁) → (Γ ⊢ t₂)  -- Applications

-- The application of the identity function to itself is a well-typed term.
example₁ : [] ⊢ ∅ ⇒ ∅
example₁ = (λx (` zero)) · (λx (` zero))
\end{code}