### 2021-03-12 paper/Paper/Section3-Kits.lagda.tex
Worked on paper.

\begin{code}[hide]
module Paper.Section3-Kits where

open import Paper.Section2-Intrinsic
open import Data.List using (List; []; _∷_)

infix  4  _∋_  _⊢_  _–[_]→_
infixr 5  _⇒_  λx_
infixl 6  _·_
\end{code}

\newpage
\section{Kits}

McBride proposed the kit technique for intrinsically typed formalizations.

\begin{code}
record Kit : Set₁ where
  infix 4 _◆_
  field
    _◆_  : List Type → Type → Set
    vr   : ∀ t → Γ ∋ t → Γ ◆ t
    tm   : ∀ t → Γ ◆ t → Γ ⊢ t
    wk   : ∀ {t'} t → Γ ◆ t → (t' ∷ Γ) ◆ t

_–[_]→_ : List Type → Kit → List Type → Set
Γ₁ –[ K ]→ Γ₂ = ∀ t → Γ₁ ∋ t → Γ₂ ◆ t where open Kit K

lift : (K : Kit) → Γ₁ –[ K ]→ Γ₂ → (t ∷ Γ₁) –[ K ]→ (t ∷ Γ₂)
lift K f _ zero    = vr _ zero     where open Kit K
lift K f _ (suc x) = wk _ (f _ x)  where open Kit K

kitᵣ : Kit
kitᵣ = record {  _◆_  =  _∋_
              ;  vr   = λ t x → x
              ;  tm   = λ t x → ` x
              ;  wk   = λ t x → suc x
              }

traverse : ∀ (K : Kit) → (Γ₁ –[ K ]→ Γ₂) → Γ₁ ⊢ t → Γ₂ ⊢ t
traverse K f (` x)      = Kit.tm K _ (f _ x)
traverse K f (λx e)     = λx (traverse K (lift K f) e)
traverse K f (e₁ · e₂)  = traverse K f e₁ · traverse K f e₂

kitₛ : Kit
kitₛ = record {  _◆_  = _⊢_
              ;  vr   = λ t x → ` x
              ;  tm   = λ t x → x
              ;  wk   = λ t x → traverse kitᵣ (Kit.wk kitᵣ) x
              }

rename : (Γ₁ –[ kitᵣ ]→ Γ₂) → Γ₁ ⊢ t → Γ₂ ⊢ t
rename ρ e = traverse kitᵣ ρ e

subst : (Γ₁ –[ kitₛ ]→ Γ₂) → Γ₁ ⊢ t → Γ₂ ⊢ t
subst σ e = traverse kitₛ σ e
\end{code}