### 2022-11-14 presentation/Presentation/Section3-Kits.lagda.tex
Update.

\begin{code}[hide]
module Presentation.Section3-Kits where

open import Paper.Section2-Intrinsic
open import Data.List using (List; []; _∷_)

infix  4  _–[_]→_
\end{code}

\section{Kits}

\begin{frame}
McBride proposed the kit technique for intrinsically typed formalizations.
\end{frame}

\begin{frame}
\begin{code}
record Kit : Set₁ where
  infix 4 _◆_
  field
    _◆_  : List Type → Type → Set
    vr   : ∀ t → (Γ ∋ t) → (Γ ◆ t)
    tm   : ∀ t → (Γ ◆ t) → (Γ ⊢ t)
    wk   : ∀ {t'} t → (Γ ◆ t) → (t' ∷ Γ ◆ t)
\end{code}
\end{frame}

\begin{frame}
\begin{code}
_–[_]→_ : List Type → Kit → List Type → Set
Γ₁ –[ K ]→ Γ₂ = ∀ t → (Γ₁ ∋ t) → (Γ₂ ◆ t) where open Kit K

lift : (K : Kit) → (Γ₁ –[ K ]→ Γ₂) → (t ∷ Γ₁ –[ K ]→ t ∷ Γ₂)
lift K f _ zero     = vr _ zero     where open Kit K
lift K f _ (suc x)  = wk _ (f _ x)  where open Kit K
\end{code}
\end{frame}

\begin{frame}
\begin{code}
kitᵣ : Kit
kitᵣ = record {  _◆_  =  _∋_
              ;  vr   = λ t x → x
              ;  tm   = λ t x → ` x
              ;  wk   = λ t x → suc x
              }
\end{code}
\end{frame}

\begin{frame}
\begin{code}
traverse : ∀ (K : Kit) → (Γ₁ –[ K ]→ Γ₂) → (Γ₁ ⊢ t) → (Γ₂ ⊢ t)
traverse K f (` x)      = Kit.tm K _ (f _ x)
traverse K f (λx e)     = λx (traverse K (lift K f) e)
traverse K f (e₁ · e₂)  = traverse K f e₁ · traverse K f e₂
\end{code}
\end{frame}

\begin{frame}
\begin{code}
kitₛ : Kit
kitₛ = record {  _◆_  = _⊢_
              ;  vr   = λ t x → ` x
              ;  tm   = λ t x → x
              ;  wk   = λ t x → traverse kitᵣ (Kit.wk kitᵣ) x
              }
\end{code}
\end{frame}

\begin{frame}
\begin{code}
rename : (Γ₁ –[ kitᵣ ]→ Γ₂) → (Γ₁ ⊢ t) → (Γ₂ ⊢ t)
rename ρ e = traverse kitᵣ ρ e

subst : (Γ₁ –[ kitₛ ]→ Γ₂) → (Γ₁ ⊢ t) → (Γ₂ ⊢ t)
subst σ e = traverse kitₛ σ e
\end{code}
\end{frame}